---
title: "BayesianRegressionUsingBGLR"
author: Tanzila
date: 5/26/2021
output: html_notebook
---

### required packages

First, load packages required for the analysis.
```{r}
require(BGLR) # package for Bayesian regression
require(here) # convert a relative path to a absolute path to a file
require(tictoc)
```
### Read data.

read genome-wide marker genotype data *only for obtaining sample names*
```{r}
c7AIR.org <- readRDS(here("C7AIR/C7AIR_Data/C7AIR_Raw", "C7AIR_Score.rds"))
```

# data preparation

read genome-wide marker genotype data.
```{r}
c7AIR.cmp <- read.csv(here("C7AIR/C7AIR_Data/C7AIR_CompressedData", "C7AIR_Compress_1_Pheno.csv"))
c7AIR.cmp <- t(c7AIR.cmp)
head(c7AIR.cmp[,1:6])
dim(c7AIR.cmp)
```
check the size of data
```{r}
dim(c7AIR.org)
dim(c7AIR.cmp)
```

check the column names.
```{r}
colnames(c7AIR.cmp)
```
```{r}
colnames(c7AIR.cmp) <- sprintf("V%05d", 1:ncol(c7AIR.cmp))
head(colnames(c7AIR.cmp))
tail(colnames(c7AIR.cmp))
```
OK!

### preparation of phenotypic data

Read phenotypic data.
```{r}
pheno <- read.csv(here("C7AIR/C7AIR_Data/C7AIR_Raw", "C7AIR_Phenotype.csv"), row.names = 1)
pheno <- na.omit(pheno)
head(pheno)
#dim(pheno)
```
read cross-validation ids.
```{r}
cv.id <- read.csv(here("C7AIR/C7AIR_Data/C7AIR_Raw", "C7AIR_Pheno_CVID.csv"), row.names = 1)
head(cv.id[,1:5])
```
```{r}
c7AIR <- c7AIR.cmp[rownames(pheno), ]
c7AIR <- c7AIR[, apply(c7AIR, 2, sd) > 0]
dim(c7AIR)
```

compare the order of rows of *pheno* with *gmat.score*.
```{r}
sum(rownames(pheno) == rownames(c7AIR))
nrow(pheno)
```

All the rows of *pheno* are arranged in the same order as *c7AIR*.

### Bayesian regression

```{r}
X <- as.matrix(c7AIR)
head(X[,1:5])
dim(X)
```

189 genotypes by 303 markers

#### Gaussian process

Gaussian process regression based on a genomic relationship matrix (a linear kernel).
```{r}
grm <- tcrossprod(scale(X)) / ncol(X)
grm[1:6, 1:6]
```
```{r}
dim(grm)
```
Perform cross-validation with Gaussian process regression.
```{r}
tic()
#X <- X[, 1:100]
y.out <- vector("list", ncol(pheno))
y.predresult <- vector("list", ncol(pheno))
names(y.out) <- colnames(pheno)
for(trait.id in 1:ncol(pheno)) {
  print(colnames(pheno)[trait.id])
  y <- pheno[, trait.id]
  names(y) <- rownames(pheno)
  y <- na.omit(y)
  this.cv.id <- cv.id[names(y), ]
  this.X <- X[names(y), ]
  y.pred <- matrix(NA, nrow(pheno), ncol(cv.id))
  rownames(y.pred) <- rownames(pheno)
  colnames(y.pred) <- colnames(cv.id)
  for(i in seq_len(ncol(this.cv.id))) {
  	for(j in 1:max(this.cv.id)) {
      print(paste(i, j))
      is.test <- this.cv.id[,i] == j
      y.train <- y
      y.train[is.test] <- NA
      eta <- list(list(K = grm, model = "RKHS"))
      #model <- BGLR(y = y.train, ETA = eta, nIter = 5000, burnIn = 1000,
      #			  verbose = F, saveAt = here("temp", "bglr_")) <- MCMC chain is too short.
      model <- BGLR(y = y.train, ETA = eta, nIter = 5000, burnIn = 1000,
					  verbose = F, saveAt = here("temp", "bglr_"))
      y.pred[names(y), i][is.test] <- model$yHat[is.test]
     }
  }
  y.out[[trait.id]] <- data.frame(pheno[, trait.id], y.pred)
  pred.acc <- cor(y.out[[trait.id]], use = "pair")[1,]
  y.predresult[[trait.id]] <- data.frame(pred.acc)
  print(cor(y.out[[trait.id]], use = "pair")[1,])
}
toc()
```
output GBLUP prediction accuracy
```{r}
write.csv(y.predresult, here("C7AIR/C7AIR_Data/C7AIR_Prediction_Accuracy",  "C7AIR_PredictionAccuracy_Compress_2_GBLUP.csv"), quote = F)
```

### Bayes B regression

```{r}
#X <- X[, 1:100]
tic()
y.out <- vector("list", ncol(pheno))
y.predresult <- vector("list", ncol(pheno))
names(y.out) <- colnames(pheno)
for(trait.id in 1:ncol(pheno)) {
  print(colnames(pheno)[trait.id])
  y <- pheno[, trait.id]
  names(y) <- rownames(pheno)
  y <- na.omit(y)
  this.cv.id <- cv.id[names(y), ]
  this.X <- X[names(y), ]
  y.pred <- matrix(NA, nrow(pheno), ncol(cv.id))
  rownames(y.pred) <- rownames(pheno)
  colnames(y.pred) <- colnames(cv.id)
  for(i in 1:ncol(this.cv.id)) {
  	for(j in 1:max(this.cv.id)) {
      print(paste(i, j))
      is.test <- this.cv.id[,i] == j
      y.train <- y
      y.train[is.test] <- NA
#     eta <- list(list(X = X, model = "BayesB")) <- X should be this.X
      eta <- list(list(X = this.X, model = "BayesB"))
      #model <- BGLR(y = y.train, ETA = eta, nIter = 5000, burnIn = 1000,
#					  verbose = F, saveAt = here("temp", "bglr_")) <- MCMC chain is too short.
      model <- BGLR(y = y.train, ETA = eta, nIter = 5000, burnIn = 1000,
					  verbose = F, saveAt = here("temp", "bglr_"))
      #y.pred[is.test] <- model$yHat[is.test]
      y.pred[names(y), i][is.test] <- model$yHat[is.test]
     }
  }
  y.out[[trait.id]] <- data.frame(pheno[, trait.id], y.pred)
  pred.acc <- cor(y.out[[trait.id]], use = "pair")[1,]
  y.predresult[[trait.id]] <- data.frame(pred.acc)
  print(cor(y.out[[trait.id]], use = "pair")[1,])
}
toc()
```

output BayesB prediction accuracy
```{r}
write.csv(y.predresult, here("C7AIR/C7AIR_Data/C7AIR_Prediction_Accuracy",  "C7AIR_PredictionAccuracy_Compress_2_BayesB.csv"), quote = F)
```
