---
title: "map marker genoetypes to phenotypes"
author: "岩田洋佳 hiroiwata@g.ecc.u-tokyo.ac.jp"
date: "2020/12/3"
output: html_document
---

required packages
```{r}
require(here)
```

read data.
```{r}
geno_1 <- read.csv(here("HDRA/HDRA_Data/HDRA_Zero_CompressedData", "HDRA_Zero_Net_1_CompressedData.csv"))
geno_2 <- read.csv(here("HDRA/HDRA_Data/HDRA_Zero_CompressedData", "HDRA_Zero_Net_2_CompressedData.csv"))
geno_3 <- read.csv(here("HDRA/HDRA_Data/HDRA_Zero_CompressedData", "HDRA_Zero_Net_3_CompressedData.csv"))
pheno <- read.csv(here("HDRA/HDRA_Data/HDRA_Raw", "RiceDiversityPheno4GWASGS.csv"), row.names = 1)
```

get column names of geno datasets.
```{r}
geno.colname <- colnames(geno_3)
```

extract required parts of IDs for marker genotype data.
```{r}
tmp <- strsplit(geno.colname, "[.]")
geno.accname <- sapply(tmp, function(x){return(x[3])})
geno.accid <- sapply(tmp, function(x){return(x[4])})
```

extract required parts of IDs for phenotype data.
```{r}
tmp <- strsplit(rownames(pheno), c("[.]"))
tmp <- sapply(tmp, function(x){return(x[1])})
tmp <- strsplit(tmp, c("@"))
pheno.accid <- sapply(tmp, function(x){return(x[2])})
```

Check the overlap between geno and pheno.
```{r}
match <- pheno.accid %in% geno.accid
sum(!match)
```
all phenotypic data are matched to the geno data.

choose accessions in pheno data from geno data and reorder them.
```{r}
geno_1_sel <- geno_1
colnames(geno_1_sel) <- geno.accid
geno_1_sel <- geno_1_sel[, pheno.accid]
 
geno_2_sel <- geno_2
colnames(geno_2_sel) <- geno.accid
geno_2_sel <- geno_2_sel[, pheno.accid]

geno_3_sel <- geno_3
colnames(geno_3_sel) <- geno.accid
geno_3_sel <- geno_3_sel[, pheno.accid]
```

check the size of selected data.
```{r}
dim(geno_1_sel)
dim(geno_2_sel)
dim(geno_3_sel)
```

output the selected and reordered data.
```{r}
write.csv(geno_1_sel, here("HDRA/HDRA_Data/HDRA_Zero_CompressedData", "HDRA_Zero_Net_1_CompressedData_Pheno.csv"), quote = F)
write.csv(geno_2_sel, here("HDRA/HDRA_Data/HDRA_Zero_CompressedData", "HDRA_Zero_Net_2_CompressedData_Pheno.csv"), quote = F)
write.csv(geno_3_sel, here("HDRA/HDRA_Data/HDRA_Zero_CompressedData", "HDRA_Zero_Net_3_CompressedData_Pheno.csv"), quote = F)
```

